---
title: "FLP-calculations"
author: "Sarah Hu"
date: "5/3/2022"
output: html_document

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Premise

Define cell concentrations
```{r}
# Based on previous Axial microbio surveys, we can assume the concentration at diffuse fluids is 1x10^5 cells/ml
insitu <- 1E5

bsw <- 1E4 #assumed in situ concentration for plume & background

# We will be collecting the resident population, concentrating it, heat-killing, staining, and then counting. 

## Our stock FLP from resident prey population (Quissett beach FLP stock)
flp_stock <- 1E8 # FLP from WHOI

flp_axial <- 6E6 #fill in with estimated stock from Axial in situ
# flp_old <- 2.79E8
```

Define experiment parameters
```{r}
# For the FLP uptake experiment set up, in ml
total_vol_SID <- 380
total_vol_1000 <- 1000
total_vol_500 <- 500
total_vol_200 <- 200

# When miniSID chamber is filled, tracer prey get added at this volume
inoc_trace_SID <- 30

# To spike bottles target making 2000 µl incoulum
inoc_trace_bottles <- 2

# Final volume of prey tracer, in ml
total_vol_tracer <- 50

## Desired final concentration of FLP
flp_final <- (0.1)*insitu
# flp_final # 10% of in situ bacteria and archaea population

flp_final_bsw <- (0.1)*bsw
# flp_final_bsw
```


## Prepare FLP working stock

When FLP working stock is 1e6, for the miniSID you end up needing ~20 ml to prep 3 bags. 
```{r}
# 1e+6
flp_working <- 1e+6 # Concentration of working stock of FLP
vol_working <- 45 #ml # for miniSID
# vol_working <- 20 #ml for shipboard
vol_stock <- (vol_working * flp_working)/flp_stock
# vol_stock <- (vol_working * flp_working)/flp_axial
vol_stock*1000 #µl

cat("To prep working stock of FLP at", vol_working, "mls, add", (vol_stock*1000), "µl of FLP stock to", (vol_working - vol_stock), "ml of buffer or ultra-cleaen seawater. Final concentration will be", flp_working, "cells/ml.")
```


### miniSID calcs

Solve for needed concentration of FLP in the FLP prey tracer bag for the miniSID
```{r}
# c1v1=c2v2
# (c1)(vol of added tracer) = (1E4)(390ml)

c1 <- (flp_final*total_vol_SID)/(inoc_trace_SID)
# set C1 to new R object
inoc_conc_SID <- c1

inoc_conc_SID # Final concentration of tracer prey bag
formatC(inoc_conc_SID, format = "e", digits = 2)
```

Now, solve for preparing the tracer prey bag.
While we assume about 30 ml of the tracer prey bag will be taken up by the main chamber, we will set the final volume of tracer prey bag to 50 ml
```{r}
#c1v2=c2v2
# (desired conc of flp in inoc)(total vol of tracer prey bag) = (stock FLP conc)(v2)
v2 <- (inoc_conc_SID*total_vol_tracer)/flp_working
# v2 <- (inoc_conc_SID*total_vol_tracer)/flp_axial
# v2
# set v1 to new R object
stock_vol_SID <- v2
stock_vol_SID # in ml, how much of stock FLP to add
stock_vol_SID*1000 # in µl
```
Prepare tracer prey bags.
```{r}
cat("Add", (stock_vol_SID*1000), "µl to", (total_vol_tracer - stock_vol_SID), "ml of ultrafiltered seawater to prep miniSID tracer prey bag for deployment at vent site. Final concentration of tracer prey in main chamber will be", (flp_final), "cells/ml.")

cat("For full miniSID deployment, prep 6 tracer prey bags with", (stock_vol_SID), "mls of working stock total. You need,", (stock_vol_SID*6), "mls total.\n\nFor 3 bags, have", (stock_vol_SID*3), "mls.\n\nFor 6 bags, have", (stock_vol_SID*6), "mls.")
```

### Shipboard

Depending on target location and available vent or sample fluid, there are a few scenarios for experiments.

#### Vent 1L 


#### Vent 500 ml

Vent 500 ml experiments have 
* 2 Control bottles (0.8µm filtered) at 250 ml total
* 3 experiment bottles at 500 ml total

Define variables for this experiment and print out set up.
```{r}
## Desired final concentration of FLP
flp_final <- (0.1)*1E5
flp_final # 10% of in situ bacteria and archaea population
ctrl_ship <- 100
# ctrl_ship <- 250

exp_ship <- 500

flp_working
flp_axial
vol_inoc_ctrl <- (flp_final * ctrl_ship)/flp_working
# vol_inoc_ctrl <- (flp_final * ctrl_ship)/flp_axial
vol_inoc_ctrl * 1000 #µl

vol_inoc_exp <- (flp_final * exp_ship)/flp_working
# vol_inoc_exp <- (flp_final * exp_ship)/flp_axial
vol_inoc_exp * 1000 #µl

cat("For shipboard vent grazing experiments where exp bottles are 500 ml and controls are 250 ml, add:\n\n", (vol_inoc_ctrl * 1000), "µl of working stock to 250 ml Control bottles\n", (vol_inoc_exp * 1000), "µl of working stock to 500 ml experimental bottles.\n\n For complete experiment with 2 controls and 3 experimental bottles, you will need", ((vol_inoc_ctrl*2) + (vol_inoc_exp*3)), "mls of FLP working stock.")
```

#### Plume & background 1000 ml

Background 1000 ml experiments have 
* 2 Control bottles (0.8µm filtered) at 500 ml total
* 3 experiment bottles at 1000 ml total

Define variables for this experiment and print out set up.
```{r}
## Desired final concentration of FLP
flp_final_bsw <- (0.1)*1E4
flp_final_bsw # 10% of in situ bacteria and archaea population

ctrl_ship <- 100 #500 ml control
# ctrl_ship <- 500 #500 ml control
exp_ship <- 1000 # 1L experimental

# flp_stock
flp_working # 1e+6

vol_inoc_ctrl <- (flp_final_bsw * ctrl_ship)/flp_working
vol_inoc_ctrl * 1000 #µl

vol_inoc_exp <- (flp_final_bsw * exp_ship)/flp_working
vol_inoc_exp * 1000 #µl


cat("For shipboard plume & background grazing experiments where exp bottles are 1000 ml and controls are 500 ml, add:\n\n", (vol_inoc_ctrl * 1000), "µl of working stock to 500 ml Control bottles\n", (vol_inoc_exp * 1000), "µl of working stock to 1000 ml experimental bottles.\n\n For complete experiment with 2 controls and 3 experimental bottles, you will need", ((vol_inoc_ctrl*2) + (vol_inoc_exp*3)), "mls of FLP working stock.")
```


## Estimating ideal stock FLP

Assumes 7 miniSID deployments and 6 shipboard experiments. 

I have been making a working stock of 1e6, and assuming the frozen stock is 1e8. This is what the FLP from WHOI is around, so we're starting here to estimate total cells needed.

Based on above, we need 38 mls of working stock (1e6) for 6 prey tracer bags for a miniSID deployment. This is already an overestimate, because we may not have that many prey tracer bags, but we aren't sure what the schedule looks like yet - so this is conservative. 

Also, based on above, for a vent shipboard experiment we need 17 mls of working stock - but I'd like to increase this in anticipation of having 1 L treatments ideally - 22 mls of working stock at 1e6. Assume I run 6 shipboard experiments. 

How many mls total of working stock do I need for all these experiments?
For miniSID, 
```{r}
# 6 bags use 38 mls * 7 deployments
38 * 7

# add 20%
((38 * 7)*(0.2)) + (38 * 7)


# 6 experiments, 22 mls each.
6 * 22

((6 * 22)*(0.2)) + (6 * 22)
```

Need a total of `319 ml` for miniSID and `158.4 mls` for shipboard.

Round up and assume we need `350 ml` of working stock at 1e6 cells/ml for the miniSID deployments.
And round up to `175 mls` for the shipboard experiment.

What about frozen stock?
```{r}
# Therefore we need
500 * 7
```

Based on above we need `~500 ul` of frozen stock to make 45 mls of working stock and `200 ul` of frozen stock to make 20 mls of working stock.
Where working stock is 1e6 and frozen stock is 1e6.

Assuming these calculations
What about frozen stock?
```{r}
# Therefore we need
min_froz <- 500 * 7 # miniSID frozen stock
min_froz

ship_froz <- 200 * 6 # shipboard frozen stock
ship_froz

total_ul <- min_froz + ship_froz
```

Together that is ```min_froz + ship_froz``` total uls of frozen stock needed at 1e8.


Estimate total cells needed:
```{r}
total_ul / 1000 # reports total ul as mls.
total_cells <- (total_ul/1000) * 1e8 #unit is cells/ml
total_cells
```
Total cells needed in the frozen stocks is ```total_cells```.

Again, to be conservative - we may lose up to 60% of the grown cells... So estimate how many cells we need with this type of error.
```{r}
# Account for this
((total_cells)*0.6) + (total_cells)
```

Can we add an order of magnitude to be safe??

*1 x 10^9*